\section{Expressing checks}

\begin{frame}[t,fragile]{How do I express contracts?}
\begin{itemize}
  \item Today operator[] is std::vector is roughly (taken for existing implementation).
\end{itemize}
\begin{block}{Current implementation}
\begin{lstlisting}
T & vector::operator[](size_t pos) noexcept
{
  return buf + pos;
}
\end{lstlisting}
\end{block}
\begin{itemize}
  \item Today this function never throws.
  \item It is marked as \cppkey{noexcept}.
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{After Tuesday}
\begin{block}{Option A}
\begin{lstlisting}
T & myvector::operator[](size_t pos) noexcept
  [[expects: 0<= pos && pos < size()]]
{
  return buf[pos];
}
\end{lstlisting}
\end{block}
\begin{itemize}
  \item If called out of contract the single global handler is called.
  \item The function is marked as \cppkey{noexcept}.
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Checking in the body}
\begin{itemize}
 \item I could still decide to have an implementation assert. 
 \item \textbf{Note}: in that case we decided that this cannot throw. 
\end{itemize}
\begin{block}{Option B}
\begin{lstlisting}
T & myvector::operator[](size_t pos) noexcept
{
  [[check: 0<= pos && pos < size()]]
  return buf[pos];
}
\end{lstlisting}
\end{block}
\begin{itemize}
  \item As we have a single global handler, even if we allowed throwing the problem
would not be solved. 
  \item Consequently the function is still noexcept.
  \item Besides static analysis would be limited for non-template non-inline.
\end{itemize}
\end{frame}

\begin{frame}[t,fragile]{Throwing an exception}
\begin{itemize}
  \item We might want to have an implementation with at()-style:
\end{itemize}
\begin{block}{Option C}
\begin{lstlisting}
T & myvector::operator[](size_t pos)
{
  if (0>pos || pos>=size())) throw out_of_range{};
  return buf[pos];
}
\end{lstlisting}
\end{block}
\begin{itemize}
  \item This would be certainly \cppkey{noexcept(false)}.
  \item But this always throws and looses the benfits from static analysis.
\end{itemize}
\end{frame}
